---
title: "Most Popular Cuisines Across the US"
author: "Kevin Chan"
date: "6 May 2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose of This Project
### The purpose of this project is to find out what types of cuisines are most popular among residents and tourists of the cities of Boston, Chicago, Dallas, and San Francisco. To do this, real-time data was obtained by tapping into the Foursquare API.

## Loading the Libraries Required for this Project
```{r}
# install.packages("rgeos")
library(rgeos)
# install.packages("rgdal")
library(rgdal)
library(httr)
library(plyr)
library(dplyr)
library(jsonlite)
library(ggplot2)
```

## The Values Needed to Tap into the Foursquare API
```{r}
client_id="GWSECN2BPBIDYSBEZD35XYTLVDYYNIHFQ3FBE2YTMO3BFJWD"
client_secret="WAZDXRFGQO12IIS00G2KZS0H34BEVDWYDDGCM4AWOWEXRF3K"
v="20180504"
query="Restaurant"
radius=10000
```

## Files Necessary to Get the Latitude and Longitude of Cities
### Note: These files were downloaded from https://developers.google.com/adwords/api/docs/appendix/geotargeting?csw=1.
```{r}
cities<-read.csv("files/cities_google.csv")
states<-read.csv("files/states_google.csv")

states<-states[,-4]
names(states)[1]<-"Parent.ID"
city.state<-merge(cities,states,by="Parent.ID",all=TRUE)
city.state<-city.state[,-c(9,10,11,12)]
names(city.state)<-c("Parent.ID","Criteria.ID","City","Canonical.Name","County.Code","Target", "Status","State")

abb<-read.csv("files/abbreviations.csv")

city_state<-merge(city.state,abb,by="State")
city_state[,10]<-as.character(city_state[,10])
city_state[,4]<-as.character(city_state[,4])
```

## Functions to Obtain the Geocode for Cities
### Note: The code is from http://stackoverflow.com/questions/27867846/quick-way-to-get-longitude-latitude-from-city-state-input
```{r}
geo_init <- function() {
  
  try({
    GET("http://www.mapcruzin.com/fcc-wireless-shapefiles/cities-towns.zip",
        write_disk("cities.zip"))
    unzip("cities.zip", exdir="cities") })
  
  shp <- readOGR("cities-towns/citiesx020.shp", "citiesx020")
  
  geo <-
    gCentroid(shp, byid=TRUE) %>%
    data.frame() %>%
    rename(lon=x, lat=y) %>%
    mutate(city=shp@data$NAME, state=shp@data$STATE)
  
}

geocode <- function(geo_db, city, state) {
  do.call(rbind.data.frame, mapply(function(x, y) {
    geo_db %>% filter(city==x, state==y)
  }, city, state, SIMPLIFY=FALSE))
}
geo_db <- geo_init()
```

## Using the Functions Above to Get the Latitude and Longitude for Boston, Chicago, Dallas, and San Francisco
```{r}
# Boston's Latitude and Longitude
boston=geo_db %>% geocode(city="Boston",state="MA")
boston_ll=paste0(boston$lat,",",boston$lon)

# Chicago's Latitude and Longitude
chicago=geo_db %>% geocode(city="Chicago",state="IL")
chicago_ll=paste0(chicago$lat,",",chicago$lon)

# Dallas' Latitude and Longitude
dallas=geo_db %>% geocode(city="Dallas",state="TX")
dallas_ll=paste0(dallas$lat,",",dallas$lon)

# San Francisco's Latitude and Longitude
sf=geo_db %>% geocode(city="San Francisco",state="CA")
sf_ll=paste0(sf$lat,",",sf$lon)
```

## The Foursquare Function that Taps into the API
```{r}
foursquare<-function(client_id,client_secret,v,query,radius,near.df){
  foursquare.list<-list();
  require(jsonlite);
  for(i in 1:length(near.df)){
    near=near.df[i];
    fqs.query=paste0("https://api.foursquare.com/v2/venues/search?client_id=",client_id,
                     "&client_secret=",client_secret,
                     "&intent=browse",
                     "&v=",v,
                     "&radius=",radius,
                     "&ll=",near,
                     "&query=",query);
    fqs.request=readLines(fqs.query);
    foursquare.list[[i]]<-fromJSON(fqs.request,simplifyDataFrame=TRUE);
    Sys.sleep(20);
  }
  return(foursquare.list);
}
```

## Calling the Foursquare API to obtain restaurant data for each of the cities
```{r}
# Boston's Restaurant Data
boston_restaurants=foursquare(client_id,client_secret,v,query,radius,boston_ll)

# Chicago's Restaurant Data
chicago_restaurants=foursquare(client_id,client_secret,v,query,radius,chicago_ll)

# Dallas' Restaurant Data
dallas_restaurants=foursquare(client_id,client_secret,v,query,radius,dallas_ll)

# San Francisco's Restaurant Data
sf_restaurants=foursquare(client_id,client_secret,v,query,radius,sf_ll)
```

## Function to Obtain List of Restaurants from Foursquare
```{r}
get_business<-function(list){
  return(list$response$venues)
}
```

## Boston's Restaurant Data
```{r}
## Obtaining list of Boston's restaurants from Foursquare.
boston_food<-lapply(boston_restaurants,get_business)
boston_test<-boston_food[sapply(boston_food, length)>=1]
boston_test<-lapply(boston_test,flatten)

boston_foodplaces<-rbind.fill(lapply(boston_test,function(y){as.data.frame((y),stringsAsFactors=F)}))
boston_foodplaces<-boston_foodplaces[!duplicated(boston_foodplaces$id),]

## Cleaning list of Boston's restaurants.
boston_temp<-data.frame(t(sapply(boston_foodplaces$categories,c)))

clean_list_data_frame<-function(data){
  for(i in 1:ncol(data)){
    data[,i]<-as.character(data[,i])
  }
  return(data)
}

boston_temp<-clean_list_data_frame(boston_temp)
boston_foodplaces<-within(boston_foodplaces,rm(categories,venueChains,location.formattedAddress,hereNow.groups,specials.items))
boston_foodplaces<-cbind(boston_foodplaces,boston_temp)
boston_foodplaces <- boston_foodplaces[,!duplicated(colnames(boston_foodplaces))]
```

## Graph of Boston's Most Popular Cuisines
```{r}
detach("package:plyr",unload=T)
boston_cuisine <- boston_foodplaces %>% group_by(shortName) %>% summarise(count = n())
ggplot(boston_cuisine, aes(x = shortName, y = count, fill = shortName)) + geom_bar(stat = "identity") + coord_flip() + 
      labs(title = "Most Popular Cuisines in Boston", x = "Cuisine", y="Frequency")
```

### Conclusion: The most popular cuisine in Boston is American cuisine, followed by Chinese cuisine. Some notes to consider are that the API collects real-time data and that it only collects a small sample of all the restaurants, so the data collected and graph are subject to change.

## Chicago's Restaurant Data
```{r}
## Obtaining list of Chicago's restaurants from Foursquare.
chicago_food<-lapply(chicago_restaurants,get_business)
chicago_test<-chicago_food[sapply(chicago_food, length)>=1]
chicago_test<-lapply(chicago_test,flatten)

library(plyr)
chicago_foodplaces<-rbind.fill(lapply(chicago_test,function(y){as.data.frame((y),stringsAsFactors=F)}))
chicago_foodplaces<-chicago_foodplaces[!duplicated(chicago_foodplaces$id),]

## Cleaning list of Chicago's restaurants.
chicago_temp<-data.frame(t(sapply(chicago_foodplaces$categories,c)))

chicago_temp<-clean_list_data_frame(chicago_temp)
chicago_foodplaces<-within(chicago_foodplaces,rm(categories,venueChains,location.formattedAddress,hereNow.groups,specials.items))
chicago_foodplaces<-cbind(chicago_foodplaces,chicago_temp)
chicago_foodplaces <- chicago_foodplaces[,!duplicated(colnames(chicago_foodplaces))]
```

## Graph of Chicago's Most Popular Cuisines
```{r}
detach("package:plyr",unload=T)
chicago_cuisine <- chicago_foodplaces %>% group_by(shortName) %>% summarise(count = n())
ggplot(chicago_cuisine, aes(x = shortName, y = count, fill = shortName)) + geom_bar(stat = "identity") + coord_flip() + 
  labs(title = "Most Popular Cuisines in Chicago", x = "Cuisine", y="Frequency")
```

### Conclusion: The most popular cuisine in Chicago is Chinese cuisine, followed by Italian and Indian cuisine as well as New American and Diner-style establishments. Some notes to consider are that the API collects real-time data and that it only collects a small sample of all the restaurants, so the data collected and graph are subject to change.

## Dallas' Restaurant Data
```{r}
## Obtaining list of Dallas' restaurants from Foursquare.
dallas_food<-lapply(dallas_restaurants,get_business)
dallas_test<-dallas_food[sapply(dallas_food, length)>=1]
dallas_test<-lapply(dallas_test,flatten)

library(plyr)
dallas_foodplaces<-rbind.fill(lapply(dallas_test,function(y){as.data.frame((y),stringsAsFactors=F)}))
dallas_foodplaces<-dallas_foodplaces[!duplicated(dallas_foodplaces$id),]
## Cleaning list of Dallas' restaurants.
dallas_temp<-data.frame(t(sapply(dallas_foodplaces$categories,c)))

dallas_temp<-clean_list_data_frame(dallas_temp)
dallas_foodplaces<-within(dallas_foodplaces,rm(categories,venueChains,location.formattedAddress,hereNow.groups,specials.items))
dallas_foodplaces<-cbind(dallas_foodplaces,chicago_temp)
dallas_foodplaces <- dallas_foodplaces[,!duplicated(colnames(dallas_foodplaces))]
```

## Graph of Dallas' Most Popular Cuisines
```{r}
detach("package:plyr", unload=T)
dallas_cuisine <- dallas_foodplaces %>% group_by(shortName) %>% summarise(count = n())
ggplot(dallas_cuisine, aes(x = shortName, y = count, fill = shortName)) + geom_bar(stat = "identity") + coord_flip() + 
  labs(title = "Most Popular Cuisines in Dallas", x = "Cuisine", y="Frequency")
```

### Conclusion: The most popular cuisine in Dallas is Chinese cuisine, followed by Italian and Indian cuisine as well as New American and Diner-style establishments. Some notes to consider are that the API collects real-time data and that it only collects a small sample of all the restaurants, so the data collected and graph are subject to change.

## San Francisco's Restaurant Data
```{r}
## Obtaining list of San Francisco's restaurants from Foursquare.
sf_food<-lapply(sf_restaurants,get_business)
sf_test<-sf_food[sapply(sf_food, length)>=1]
sf_test<-lapply(sf_test,flatten)

library(plyr)
sf_foodplaces<-rbind.fill(lapply(sf_test,function(y){as.data.frame((y),stringsAsFactors=F)}))
sf_foodplaces<-sf_foodplaces[!duplicated(sf_foodplaces$id),]

## Cleaning list of San Francisco's restaurants.
sf_temp<-data.frame(t(sapply(sf_foodplaces$categories,c)))

sf_temp<-clean_list_data_frame(sf_temp)
sf_foodplaces<-within(sf_foodplaces,rm(categories,venueChains,location.formattedAddress,hereNow.groups,specials.items))
sf_foodplaces<-cbind(sf_foodplaces,chicago_temp)
sf_foodplaces <- sf_foodplaces[,!duplicated(colnames(sf_foodplaces))]
```

## Graph of San Francisco's Most Popular Cuisines
```{r}
detach("package:plyr",unload=T)
sf_cuisine <- sf_foodplaces %>% group_by(shortName) %>% summarise(count = n())
ggplot(sf_cuisine, aes(x = shortName, y = count, fill = shortName)) + geom_bar(stat = "identity") + coord_flip() + 
  labs(title = "Most Popular Cuisines in San Francisco", x = "Cuisine", y="Frequency")
```

### Conclusion: The most popular cuisine in San Francisco is Chinese cuisine, followed by Italian and Indian cuisine as well as New American and Diner-style establishments. Some notes to consider are that the API collects real-time data and that it only collects a small sample of all the restaurants, so the data collected and graph are subject to change.